### 2007-01-09 Documentation/part1.tex
*** empty log message ***

TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
CHANGES ETC. MAYBE ONLY COMMENT ABOUT THAT [WW].

@@

Assign values to our standard oscillation parameters:

@@

 glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm);

@@

only $\deltacp$ is a free parameter:

@@

@@

@@

@@

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{norm}{C-Code}
\index{norm}{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NFstandard.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$
as well as some other information, the parameter type \GLB{\_params} is used. In general,
 this type is often
transferred to and from \GLOBES\ functions. Therefore, the memory 
for these vectors has to be reserved (allocated) before they can be used, 
and it has
to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
\index{norm}{Matter density!profile} 
which is stored in the density information of {\tt glb\_params}. 
Each of
these scaling factors has $1.0$ as pre-defined value. Since it
is in most cases not necessary to change this value, the user does
not need to take care of it. For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, freed and allocated again.

Another piece of information will be returned from the minimizers 
(\cf, \Chapt~\ref{chapt:correlations}) and transferred
into the {\tt glb\_params} structure is the number
of iterations used for the minimization, which is proportional to the
running time of the minimizer. In general, the user does not need to
access the elements in {\tt glb\_params} directly. A number of functions
is provided to handle these parameter structures:
\begin{function}
\GLBNS{AllocParams}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it.
\end{function}
\begin{function}
\GLBNS{AllocParams}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{DefineParams}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dm21, double dm31)} assigns
the complete set of oscillation parameters to the vector {\tt in}, which has
to be allocated before. The return value is the pointer to {\tt in} if 
the assignment was successful, and {\tt NULL} otherwise.
\end{function}
\begin{function}
 \GLBNS{CopyParams}
 {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
 copies the vector {\tt source} to the vector {\tt destination}. The return
 value is {\tt NULL} if the assignment was not successful.
\end{function}
\begin{function}
\GLBNS{PrintParams}
{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
parameters in {\tt in} to the file {\tt stream}. The oscillation
parameters, all density values, and the number of iterations are
printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
to print to the screen.
\end{function}
In addition to these basic functions, there are functions to access
the individual parameters within the parameter vectors:
\begin{function}
\GLBNS{SetOscParams}
{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
sets the oscillation parameter {\tt which} in the structure {\tt in}
to the value {\tt osc}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetOscParams}
{\tt double glbGetOscParams(glb\_params in, int which)}
returns the value of the oscillation parameter {\tt which} in the 
structure {\tt in}.
\end{function}
In both of these functions, the parameter {\tt which} runs from
$0$ to $5$ (or the number of oscillation parameters-1), where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
\GLBC{GLB\_DM\_SOL}, or \GLBC{GLB\_DM\_ATM} can be used.

@@

In order to keep track of the used version of \GLOBES , the software
provides a number of functions to check the \GLOBES\ and experiment
versions. It is up to the user to implement mechanisms into the
program and \AEDL\ files to check whether
\begin{itemize}
\item
 The program should only run with this specific version of \GLOBES .
\item
 The program can only run up to a certain \GLOBES\ version.
\item
 The program can only run with a minimum version of \GLOBES .
\item
 The program and AEDL file versions are compatible.
\end{itemize}
However, note that GLOBES~3.0 and higher requires that the \GLOBES\ version be at least as new as the
version the AEDL file was written for.
%
The functions in \GLOBES\ for version control are:
\begin{function}
\GLBNS{TestReleaseVersion}
{\tt int glbTestReleaseVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer. 
\end{function}
\begin{function}
\GLBNS{TestLibraryVersion}
{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer.
Note that the library and \GLOBES\ versions are not the same. 
\end{function}
WHAT IS THE DIFFERENCE BETWEEN THE LIBRARY AND GLOBES VERSION???? [WW]
\begin{function}
\GLBNS{VersionOfExperiment} 
{\tt const char* glbVersionOfExperiment(int experiment)} returns
the version string of the experiment number {\tt experiment} (set by
{\tt \$version} in AEDL).
The version string is allocated within the experiment structure, which means
that it cannot be altered and must not be freed by the
user.
\end{function}

@@

\example{Correlation between $\stheta$ and $\deltacp$}{
\label{ex:corrth13dcp}
\index{norm}{Correlation!two-parameter}
A typical and fast application for \GLB{ChiSys} is the visualization of
 two-parameter correlations using systematics only. For example, to
  calculate the two-parameter 
 correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
 can use the following code excerpt from {\tt example1.c}:
\begin{quote}
{\tt {\footnotesize
  /* Initialize parameter vector(s) and compute simulated data */ \\
  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
  glbDefineParams(test\_values,theta12,theta13,theta23,deltacp,sdm,ldm);  \\
  glbSetOscillationParameters(true\_values); glbSetRates(); \\
\\
  /* Iteration over all values to be computed */ \\
  double x,y,res; \\   
  for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
  for(y=0.0;y<200.0+0.01;y=y+200.0/50) \\
  \{ \\
\hspace*{0.5cm} /* Set parameters in vector of test values */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,GLB\_THETA\_13); \\
\hspace*{0.5cm} glbSetOscParams(test\_values,y*M\_PI/180.0,GLB\_DELTA\_CP); \\
  \\
\hspace*{0.5cm} /* Compute Chi2 for all loaded experiments and all rules */ \\
\hspace*{0.5cm} res=glbChiSys(test\_values,GLB\_ALL,GLB\_ALL); \\
\hspace*{0.5cm} AddToOutput(x,y,res); \\
  \} \\
} }
\end{quote}

@@

{\tt int glbDefineChiFunction(glb\_chi\_function chi\_func, int dim, const char *name)}

@@

identified by the string {\tt name} with {\tt dim} systematics parameters. Returns zero, if successful.

@@

{\tt double (*glb\_chi\_function)(int exp, int rule, int dim, double *params, double *errors)}.
In this function type, {\tt exp} is the experiment number, {\tt rule} is the rule number, {\tt dim} is the number of systematics parameters, {\tt params} is an array of the systematics auxiliary parameters themselves, and {\tt errors} is an array with the systematical errors. Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices are running from $0$ to {\tt dim}-1.

@@

Let us now discuss a simple example.  Typical applications for user-defined systematics are reactor neutrino experiments, where systematics play a crucial role. For a setup with near and far detectors,
the formula for $\chi^2$ is
\begin{align}
  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
\label{equ:chi2reactor}
\end{align}
where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
far detector, calculated for the assumed ``true'' values of the oscillation parameters.
$T_{d,i}$ are the event rates for the parameter values that are currently being tested.
$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
mass of the two detectors. In this example their central values are assumed to be zero,
while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
%
The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
the Poisson distribution, while the terms in the second line are penalties
for deviations of the systematics parameters from their central values. 
%
We use two AEDL files for this experiment, one for the far detector and one for the
near detector.
The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
\hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\
> 
}
\end{quote}
In this case, the systematics is called ``{\tt chiDCNorm}'' for systematics on, whereas systematics
off is computed with spectral information, but without systematical errors.
For the near detector AEDL-file, we have instead
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiZero" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ \} \\
> 
}
\end{quote}
In this case, the systematics {\tt chiZero} is used for systematics on, which means that there will
be no active $\chi^2$ calculation in this rule. With this definition, the user-defined systematics
will only be called once for the far detector. However, the rates from the near detector will be 
passively provided for the common $\chi^2$ function.
You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb}, you will need for {\tt example5.c}, in the example directory. 
%
We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
where we in addition include an uncorrelated energy calibration error.
MAYBE: ADDITIONAL PLOT TO SHOW RESULT? [WW]

@@

This examples demonstrates how to use a non-Gaussian error for the
external solar mixing angle input instead of a Gaussian error.
The user-defined prior function reads as follows and is very similar to the standard prior:
\begin{quote}
{\tt {\footnotesize
 double my\_prior(const glb\_params in) \\
\{ \\
  \hspace*{0.5cm} glb\_params central\_values = glbAllocParams(); \\
  \hspace*{0.5cm} glb\_params input\_errors = glbAllocParams(); \\
  \hspace*{0.5cm}  glb\_projection p = glbAllocProjection(); \\
  \hspace*{0.5cm}  glbGetCentralValues(central\_values); \\
  \hspace*{0.5cm}  glbGetInputErrors(input\_errors); \\
  \hspace*{0.5cm}  glbGetProjection(p); \\
  \hspace*{0.5cm}  int i; \\
  \hspace*{0.5cm}  double pv = 0.0; \\
  \hspace*{0.5cm}  double fitvalue,centralvalue,inputerror;

@@

THIS CHAPTER MAY NEED SOME MODIFICATIONS, SINCENOT YET FINALLY IMPLEMENTED! [WW]

@@

\section{Baseline and matter density profile}
\index{norm}{Baseline!change}
\index{norm}{Matter density!change profile}

@@

TO BE EXTENDED BY LIST HANDLING (TRANSFER LISTS IN AEDL VARIABLES). STATUS? [WW].

>>>>

TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
CHANGES ETC. ALSO EXPERIMENT (OLD) NUFACT DOES NOT EXIST ANYMORE. AND: SOME ERROR IN THIS OLD
NUFACT-FILE [WW].

CHECK: GLBSETPROJECTION REALLY USED WITH GLB\_FIXED FOR DENSITY PARAMS? IS NFSTANDARD USED?

@@

Assign values to our standard oscillation parameters and the standard matter density scaling factors:

@@

 glbSetDensityParams(true\_values,1.0,GLB\_ALL);

@@

only $\deltacp$ (and the matter density) is a free parameter:

@@

  glbSetDensityProjectionFlag(myprojection,GLB\_FREE,GLB\_ALL); \\

@@

  glbSetDensityParams(starting\_errors,1.0,GLB\_ALL); \\

@@

  glbSetDensityParams(true\_values,1.0,GLB\_ALL); \\

@@

  glbSetDensityParams(starting\_values,1.0,GLB\_ALL); \\

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{norm}{C-Code}
\index{norm}{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NFstandard.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\
    \hspace*{0.5cm}
     glbSetDensityParams(true\_values,1.0,GLB\_ALL); // Matter scaling \\

@@

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$
as well as some other information, the parameter type \GLB{\_params} is used. In general,
 this type is often
transferred to and from \GLOBES\ functions. Therefore, the memory 
for these vectors has to be reserved (allocated) before they can be used, 
and it has
to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
\index{norm}{Matter density!profile} 
which is stored in the density information of {\tt glb\_params}.
% CHANGED IN GLOBES 3.0: 
% Each of
% these scaling factors has $1.0$ as pre-defined value. Since it
% is in most cases not necessary to change this value, the user does
% not need to take care of it. 
For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, freed and allocated again.

Another piece of information will be returned from the minimizers 
(\cf, \Chapt~\ref{chapt:correlations}) and transferred
into the {\tt glb\_params} structure is the number
of iterations used for the minimization, which is proportional to the
running time of the minimizer. In general, the user does not need to
access the elements in {\tt glb\_params} directly. A number of functions
is provided to handle these parameter structures:
\begin{function}
\GLBNS{AllocParams}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it. All parameter values are
initially set to {\tt nan} (not a number).
\end{function}
\begin{function}
\GLBNS{AllocParams}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{DefineParams}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dm21, double dm31)} assigns
the complete set of oscillation parameters to the vector {\tt in}, which has
to be allocated before. The return value is the pointer to {\tt in} if 
the assignment was successful, and {\tt NULL} otherwise.
\end{function}
\begin{function}
 \GLBNS{CopyParams}
 {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
 copies the vector {\tt source} to the vector {\tt destination}. The return
 value is {\tt NULL} if the assignment was not successful.
\end{function}
\begin{function}
\GLBNS{PrintParams}
{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
parameters in {\tt in} to the file {\tt stream}. The oscillation
parameters, all density values, and the number of iterations are
printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
to print to the screen.
\end{function}
In addition to these basic functions, there are functions to access
the individual parameters within the parameter vectors:
\begin{function}
\GLBNS{SetOscParams}
{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
sets the oscillation parameter {\tt which} in the structure {\tt in}
to the value {\tt osc}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetOscParams}
{\tt double glbGetOscParams(glb\_params in, int which)}
returns the value of the oscillation parameter {\tt which} in the 
structure {\tt in}.
\end{function}
In both of these functions, the parameter {\tt which} runs from
$0$ to $5$ (or the number of oscillation parameters-1), where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
\GLBC{GLB\_DM\_SOL}, or \GLBC{GLB\_DM\_ATM} can be used.

@@

In order to keep track of the used version of \GLOBES , the software
provides a number of functions to check the \GLOBES\ and experiment
versions. It is up to the user to implement mechanisms into the
program and \AEDL\ files to check whether
\begin{itemize}
\item
 The program should only run with this specific version of \GLOBES .
\item
 The program can only run up to a certain \GLOBES\ version.
\item
 The program can only run with a minimum version of \GLOBES .
\item
 The program and AEDL file versions are compatible.
\end{itemize}
However, note that GLOBES~3.0 and higher requires that the \GLOBES\ version be at least as new as the
version the AEDL file was written for.
%
The functions in \GLOBES\ for version control are:
\begin{function}
\GLBNS{TestReleaseVersion}
{\tt int glbTestReleaseVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer. 
\end{function}
% UNDOCUMENTED FUNCTION, SINCE VERY DIFICULT TO EXPLAIN DIFFERENCE:
% \begin{function}
% \GLBNS{TestLibraryVersion}
% {\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
% the {\tt version} string of the format ``X.Y.Z'' is exactly the
% used \GLOBES\ library version, $1$ if it is older, and $-1$ if it is newer.
% \end{function}
% Note that the library and \GLOBES\ versions are not the same.

\begin{function}
\GLBNS{VersionOfExperiment} 
{\tt const char* glbVersionOfExperiment(int experiment)} returns
the version string of the experiment number {\tt experiment} (set by
{\tt \$version} in AEDL).
The version string is allocated within the experiment structure, which means
that it cannot be altered and must not be freed by the
user.
\end{function}

@@

\example{Correlation between $\stheta$ and $\deltacp$}{
\label{ex:corrth13dcp}
\index{norm}{Correlation!two-parameter}
A typical and fast application for \GLB{ChiSys} is the visualization of
 two-parameter correlations using systematics only. For example, to
  calculate the two-parameter 
 correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
 can use the following code excerpt from {\tt example1.c}:
\begin{quote}
{\tt {\footnotesize
  /* Initialize parameter vector(s) and compute simulated data */ \\
  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
  glbSetDensityParams(true\_values,1.0,GLB\_ALL); \\
  glbCopyParams(true\_values,test\_values);  \\
  glbSetOscillationParameters(true\_values); glbSetRates(); \\
\\
  /* Iteration over all values to be computed */ \\
%  double x,y,res; \\   
  for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
  for(y=0.0;y<200.0+0.01;y=y+200.0/50) \\
  \{ \\
\hspace*{0.5cm} /* Set parameters in vector of test values */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,GLB\_THETA\_13); \\
\hspace*{0.5cm} glbSetOscParams(test\_values,y*M\_PI/180.0,GLB\_DELTA\_CP); \\
  \\
\hspace*{0.5cm} /* Compute Chi2 for all loaded experiments and all rules */ \\
\hspace*{0.5cm} res=glbChiSys(test\_values,GLB\_ALL,GLB\_ALL); \\
\hspace*{0.5cm} AddToOutput(x,y,res); \\
  \} \\
} }
\end{quote}

@@

{\tt int glbDefineChiFunction(glb\_chi\_function chi\_func, int dim, const char *name, void* user\_data)}

@@

identified by the string {\tt name} with {\tt dim} systematics parameters. The parameter {\tt user\_data}
is an arbitrary pointer being transferred to {\tt chi\_func}. It can, for instance, be used to avoid
global variables. Returns zero, if successful.

@@

{\tt double (*glb\_chi\_function)(int exp, int rule, int dim, double *params, double *errors, void* user\_data)}.
In this function type, {\tt exp} is the experiment number, {\tt rule} is the rule number, {\tt dim} is the number of systematics parameters, {\tt params} is an array of the systematics auxiliary parameters themselves, and {\tt errors} is an array with the systematical errors. The parameter {\tt user\_data} is set as defined with
{\tt glbDefineChiFunction}.
Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices are running from $0$ to {\tt dim}-1.

@@

Let us now discuss a simple example.  Typical applications for user-defined systematics are reactor neutrino experiments, where systematics play a crucial role. For a setup with near and far detectors,
the formula for $\chi^2$ is
\begin{align}
  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
\label{equ:chi2reactor}
\end{align}
where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
far detector, calculated for the assumed ``true'' values of the oscillation parameters.
$T_{d,i}$ are the event rates for the parameter values that are currently being tested.
$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
mass of the two detectors. In this example their central values are assumed to be zero,
while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
%
The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
the Poisson distribution, while the terms in the second line are penalties
for deviations of the systematics parameters from their central values. 
%
We use two AEDL files for this experiment, one for the far detector and one for the
near detector.
The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
\hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\
> 
}
\end{quote}
In this case, the systematics is called ``{\tt chiDCNorm}'' for systematics on, whereas systematics
off is computed with spectral information, but without systematical errors.
For the near detector AEDL-file, we have instead
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiZero" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ \} \\
> 
}
\end{quote}
In this case, the systematics {\tt chiZero} is used for systematics on, which means that there will
be no active $\chi^2$ calculation in this rule. With this definition, the user-defined systematics
will only be called once for the far detector. However, the rates from the near detector will be 
passively provided for the common $\chi^2$ function.
You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb}, you will need for {\tt example5.c}, in the example directory. 
%
We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
where we in addition include an uncorrelated energy calibration error. See \figu{reactor} for the result
of this example.
%
\begin{figure}[t]
\begin{center}
\includegraphics[width=10cm]{reactor}
\end{center}
\caption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
\end{figure}

@@

This examples demonstrates how to use a non-Gaussian error for the
external solar mixing angle input instead of a Gaussian error.
The user-defined prior function reads as follows and is very similar to the standard prior:
\begin{quote}
{\tt {\footnotesize
 double my\_prior(const glb\_params in, void* user\_data) \\
\{ \\
  \hspace*{0.5cm} glb\_params central\_values = glbAllocParams(); \\
  \hspace*{0.5cm} glb\_params input\_errors = glbAllocParams(); \\
  \hspace*{0.5cm}  glb\_projection p = glbAllocProjection(); \\
  \hspace*{0.5cm}  glbGetCentralValues(central\_values); \\
  \hspace*{0.5cm}  glbGetInputErrors(input\_errors); \\
  \hspace*{0.5cm}  glbGetProjection(p); \\
  \hspace*{0.5cm}  int i; \\
  \hspace*{0.5cm}  double pv = 0.0; \\
  \hspace*{0.5cm}  double fitvalue,centralvalue,inputerror;

@@

@@

\section{Baseline and matter density profile}
\label{sec:baselinemd}
\index{norm}{Baseline!change}
\index{norm}{Matter density!change profile}

@@
